
#  **JAGUAR DATASET - PANTANAL  (Sao  Bento)**

#### *Alan E. de Barros, Bernardo Niebuhr, Vanesa Bejarano, Julia Oshima,Claudia Kanda, Milton Ribeiro, Ronaldo Morato,Paulo Prado*
date: "April, 04 2019"
##### Scripts adapted from Johannes Signer and John John Fieberg's lectures.

#### Run JaguarDataPrep first !!! 

```{r }
source("JaguarDataPrep.R")
```


### Preliminary tests for SSF using data from Sao Bento, Pantanal 

```{r }
options(width=165,digits.secs = 3)
opts_chunk$set(fig.width=12,fig.height=4.5, error=TRUE,cache = F)
```

Record time for running all code

```{r }
ptm<-proc.time()
```

Set the seed for the random number generator, so it will be possible
to reproduce the random points

```{r }
set.seed(10299)
options(max.print=1000000)

SaoBentotrk -> trk	
trk # %>% print(n = 9000)
summary(trk)
colSums(ifelse(table(trk$id,trk$sex)>0,1,0)) #table(droplevels(trk$id),trk$sex)  # sex
colSums(ifelse(table(trk$id,trk$age)>0,1,0)) #table(droplevels(trk$id),trk$age)  # age
colSums(ifelse(table(trk$id,trk$weight)>0,1,0)) #table(droplevels(trk$id),trk$weight) # weight
colSums(ifelse(table(trk$id,trk$status)>0,1,0)) #table(droplevels(trk$id),trk$status) # residence status
```

### Visualize all data with leaflet andd plots

```{r }
pal <- colorFactor(palette = 'Paired',domain = SaoBento$id)
leaflet(SaoBento)%>%addTiles()%>%addCircles(SaoBento$x,SaoBento$y,color=~pal(id))%>% addProviderTiles(providers$Esri.WorldImagery)
```

##### 3D plot

```{r }
open3d()
# To get a bigger window than the default
par3d(windowRect = c(100, 100, 612, 612))
Sys.sleep(0.1) # Allow sluggish window managers to catch up
with(SaoBento, plot3d(x,y,date, type="l", col=as.favtor(SaoBento$id)))
(stcube<-with(SaoBento, plot3d(x,y,date, type="l",col=as.numeric(cut(SaoBento$weight,11)), alpha=0.4)))
# Or with points
open3d()
# To get a bigger window than the default
par3d(windowRect = c(100, 100, 612, 612))
Sys.sleep(0.1) # Allow sluggish window managers to catch up

with(SaoBento, plot3d(x,y,date, type="p", col=as.favtor(SaoBento$id)))
(stcube<-with(SaoBento, plot3d(x,y,date, type="p",col=as.numeric(cut(SaoBento$weight,11)), alpha=0.4)))
```

## Some plots of movement characteristics

We can select id and steps, unnest the new data_frame and create a plot of the step-length distributions.

```{r }
trk %>% select(id, sl) %>% unnest %>% 
  ggplot(aes(sl, fill = factor(id))) + geom_density(alpha = 0.4)
```

### Absolute angles (for each movement) relative to North 
We could use a rose diagram (below) to depict the distribution of angles. 

```{r fig.height=12, fig.width=12}
ggplot(trk, aes(x = dir_abs, y=..density..)) + geom_histogram(breaks = seq(0,360, by=20))+
  coord_polar(start = 0) + theme_minimal() + 
  scale_fill_brewer() + ylab("Density") + ggtitle("Angles Direct") + 
  scale_x_continuous("", limits = c(0, 360), breaks = seq(0, 360, by=20), 
                     labels = seq(0, 360, by=20))+
  facet_wrap(~id)
```

### Turning angles 

Note: a 0 indicates the animal continued to move in a straight line, a 180 
indicates the animal turned around (but note, resting + measurement error often can
make it look like the animal turned around).

```{r fig.height=12, fig.width=12}
ggplot(trk, aes(x = dir_rel, y=..density..)) + geom_histogram(breaks = seq(-180,180, by=20))+
  coord_polar(start = 0) + theme_minimal() + 
  scale_fill_brewer() + ylab("Density") + ggtitle("Angles Direct") + 
  scale_x_continuous("", limits = c(-180, 180), breaks = seq(-180, 180, by=20), 
                     labels = seq(-180, 180, by=20))+
  facet_wrap(~id)
```

### Turning angles as histograms

```{r fig.height=12, fig.width=12}
ggplot(trk, aes(x = dir_rel)) +  geom_histogram(breaks = seq(-180,180, by=20))+
  theme_minimal() + 
  scale_fill_brewer() + ylab("Count") + ggtitle("Angles Relative") + 
  scale_x_continuous("", limits = c(-180, 180), breaks = seq(-180, 180, by=20),
                     labels = seq(-180, 180, by=20))+facet_wrap(~id, scales="free")
```

### Net-squared displacement over time for each individual

```{r fig.height=12, fig.width=12}
ggplot(trk, aes(x = t_, y=nsd_)) + geom_point()+
  facet_wrap(~id, scales="free")
```

## Explore movement characteristics by (day/night, hour, month)

### step length distribution by day/night


```{r fig.height=12, fig.width=12, warning=FALSE, message=FALSE}
ggplot(trk, aes(x = tod_, y = log(sl)))+ geom_boxplot()+geom_smooth()+facet_wrap(~id)
```

## GIS LAYERS in UTM  **These NEED BE ADJUSTED TO EACH INDIVIDUAL!!!**

```{r }
#  For now we use bases from ID 115 because its 70 Km radius emcompasses all the other individuals
```


Distances **These NEED BE RE-CALCULATED!!!**

```{r }
dist2drainage <-"D:/GISUTM/J115/UTM_b115_dist2drainage_exp.tif"
dist2water<-"D:/GISUTM/J115/UTM_b115_dist2waterbodies_exp.tif"
dist2forest <-"D:/GISUTM/J115/UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist.tif"
dist2roads <-"D:/GISUTM/J115/UTM_b115_dist2roads_exp.tif"
```

Anthropic

```{r }
human_footprint<-"D:/GISUTM/J115/UTM_b115_human_footprint_2009_1km_tif_exp.tif"
Livestock<-"D:/GISUTM/J115/UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp.tif"
popdensity<-"D:/GISUTM/J115/UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp.tif"
```

Landscape classificatiom and physical

```{r }
Landcover <-"D:/GISUTM/J115/UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp.tif"
treecover <-"D:/GISUTM/J115/UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp.tif"
topo<-"D:/GISUTM/J115/UTM_b115_Neotropic_Earthenv_dem90m_tif_exp.tif"
water<-"D:/GISUTM/J115/UTM_b115_water_frequency_2010_30m_tif_exp.tif"
```

Check projection and images

```{r }
dist2drainage=raster(dist2drainage);crs(dist2drainage); plot(dist2drainage)
dist2water=raster(dist2water); crs(dist2water); plot(dist2water)
dist2forest=raster(dist2forest); crs(dist2forest); plot(dist2forest)
dist2roads=raster(dist2roads); crs(dist2roads); plot(dist2roads)
human_footprint=raster(human_footprint); crs(human_footprint); plot(human_footprint)
Livestock=raster(Livestock); crs(Livestock); plot(Livestock)
popdensity=raster(popdensity); crs(popdensity); plot(popdensity)
Landcover=raster(Landcover); crs(Landcover); plot(Landcover)
treecover=raster(treecover); crs(treecover); plot(treecover)
topo=raster(topo); crs(topo); plot(topo)
water=raster(water); crs(water); plot(water)
```


## SSF prep


```{r }
SaoBentotrk -> trk	
```

SSFs assume that data have been collected at regular time intervals.
We can use the track_resample function to regularize the trajectory so that
all points are located within some tolerence of each other in time. To figure
out a meaningful tolerance range, we should calculate time differences between
locations & look at as a function of individual.

```{r }
(timestats<-trk %>% nest(-id,-sex,-age,-weight,-status) %>% mutate(sr = map(data, summarize_sampling_rate, time_unit = "hour")) %>%
    select(id,sex,age,weight,status,sr) %>% unnest)
	
## Lets add on the time difference to each obs.  Alerady DONE!!! 
#trk<-trk %>% group_by(id) %>% mutate(dt_ = t_ - lag(t_, default = NA))
#trk
```

Jaguar 105
Let's illustrate track regularization with ID = 105. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp105<-trk %>% filter(id=="105") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp105 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp105)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf105 <- temp105 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf105 <- ssf105 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf105 <- ssf105 %>% extract_covariates(dist2drainage); ssf105 <- rename(ssf105, dist2drainage = UTM_b115_dist2drainage_exp)
ssf105 <- ssf105 %>% extract_covariates(dist2water); ssf105 <- rename(ssf105, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf105 <- ssf105 %>% extract_covariates(dist2forest);ssf105 <- rename(ssf105, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf105 <- ssf105 %>% extract_covariates(dist2roads); ssf105 <- rename(ssf105, dist2roads = UTM_b115_dist2roads_exp) 
ssf105 <- ssf105 %>% extract_covariates(human_footprint); ssf105 <- rename(ssf105, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf105 <- ssf105 %>% extract_covariates(Livestock); ssf105 <- rename(ssf105, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf105 <- ssf105 %>% extract_covariates(popdensity); ssf105 <- rename(ssf105, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf105 <- ssf105 %>% extract_covariates(Landcover); ssf105 <- rename(ssf105, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf105 <- ssf105 %>% extract_covariates(treecover); ssf105 <- rename(ssf105, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf105 <- ssf105 %>% extract_covariates(topo); ssf105 <- rename(ssf105, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf105 <- ssf105 %>% extract_covariates(water); ssf105 <- rename(ssf105, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf105, width=Inf)
```


Jaguar 106
Let's illustrate track regularization with ID = 106. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp106<-trk %>% filter(id=="106") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp106 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp106)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf106 <- temp106 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf106 <- ssf106 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf106 <- ssf106 %>% extract_covariates(dist2drainage); ssf106 <- rename(ssf106, dist2drainage = UTM_b115_dist2drainage_exp)
ssf106 <- ssf106 %>% extract_covariates(dist2water); ssf106 <- rename(ssf106, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf106 <- ssf106 %>% extract_covariates(dist2forest);ssf106 <- rename(ssf106, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf106 <- ssf106 %>% extract_covariates(dist2roads); ssf106 <- rename(ssf106, dist2roads = UTM_b115_dist2roads_exp) 
ssf106 <- ssf106 %>% extract_covariates(human_footprint); ssf106 <- rename(ssf106, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf106 <- ssf106 %>% extract_covariates(Livestock); ssf106 <- rename(ssf106, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf106 <- ssf106 %>% extract_covariates(popdensity); ssf106 <- rename(ssf106, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf106 <- ssf106 %>% extract_covariates(Landcover); ssf106 <- rename(ssf106, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf106 <- ssf106 %>% extract_covariates(treecover); ssf106 <- rename(ssf106, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf106 <- ssf106 %>% extract_covariates(topo); ssf106 <- rename(ssf106, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf106 <- ssf106 %>% extract_covariates(water); ssf106 <- rename(ssf106, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf106, width=Inf)
```


Jaguar 107
Let's illustrate track regularization with ID = 107. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp107<-trk %>% filter(id=="107") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp107 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp107)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf107 <- temp107 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf107 <- ssf107 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf107 <- ssf107 %>% extract_covariates(dist2drainage); ssf107 <- rename(ssf107, dist2drainage = UTM_b115_dist2drainage_exp)
ssf107 <- ssf107 %>% extract_covariates(dist2water); ssf107 <- rename(ssf107, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf107 <- ssf107 %>% extract_covariates(dist2forest);ssf107 <- rename(ssf107, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf107 <- ssf107 %>% extract_covariates(dist2roads); ssf107 <- rename(ssf107, dist2roads = UTM_b115_dist2roads_exp) 
ssf107 <- ssf107 %>% extract_covariates(human_footprint); ssf107 <- rename(ssf107, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf107 <- ssf107 %>% extract_covariates(Livestock); ssf107 <- rename(ssf107, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf107 <- ssf107 %>% extract_covariates(popdensity); ssf107 <- rename(ssf107, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf107 <- ssf107 %>% extract_covariates(Landcover); ssf107 <- rename(ssf107, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf107 <- ssf107 %>% extract_covariates(treecover); ssf107 <- rename(ssf107, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf107 <- ssf107 %>% extract_covariates(topo); ssf107 <- rename(ssf107, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf107 <- ssf107 %>% extract_covariates(water); ssf107 <- rename(ssf107, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf107, width=Inf)
```


Jaguar 108
Let's illustrate track regularization with ID = 108. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp108<-trk %>% filter(id=="108") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp108 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp108)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf108 <- temp108 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf108 <- ssf108 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf108 <- ssf108 %>% extract_covariates(dist2drainage); ssf108 <- rename(ssf108, dist2drainage = UTM_b115_dist2drainage_exp)
ssf108 <- ssf108 %>% extract_covariates(dist2water); ssf108 <- rename(ssf108, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf108 <- ssf108 %>% extract_covariates(dist2forest);ssf108 <- rename(ssf108, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf108 <- ssf108 %>% extract_covariates(dist2roads); ssf108 <- rename(ssf108, dist2roads = UTM_b115_dist2roads_exp) 
ssf108 <- ssf108 %>% extract_covariates(human_footprint); ssf108 <- rename(ssf108, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf108 <- ssf108 %>% extract_covariates(Livestock); ssf108 <- rename(ssf108, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf108 <- ssf108 %>% extract_covariates(popdensity); ssf108 <- rename(ssf108, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf108 <- ssf108 %>% extract_covariates(Landcover); ssf108 <- rename(ssf108, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf108 <- ssf108 %>% extract_covariates(treecover); ssf108 <- rename(ssf108, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf108 <- ssf108 %>% extract_covariates(topo); ssf108 <- rename(ssf108, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf108 <- ssf108 %>% extract_covariates(water); ssf108 <- rename(ssf108, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf108, width=Inf)
```


Jaguar 109
Let's illustrate track regularization with ID = 109. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp109<-trk %>% filter(id=="109") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp109 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp109)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf109 <- temp109 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf109 <- ssf109 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf109 <- ssf109 %>% extract_covariates(dist2drainage); ssf109 <- rename(ssf109, dist2drainage = UTM_b115_dist2drainage_exp)
ssf109 <- ssf109 %>% extract_covariates(dist2water); ssf109 <- rename(ssf109, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf109 <- ssf109 %>% extract_covariates(dist2forest);ssf109 <- rename(ssf109, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf109 <- ssf109 %>% extract_covariates(dist2roads); ssf109 <- rename(ssf109, dist2roads = UTM_b115_dist2roads_exp) 
ssf109 <- ssf109 %>% extract_covariates(human_footprint); ssf109 <- rename(ssf109, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf109 <- ssf109 %>% extract_covariates(Livestock); ssf109 <- rename(ssf109, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf109 <- ssf109 %>% extract_covariates(popdensity); ssf109 <- rename(ssf109, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf109 <- ssf109 %>% extract_covariates(Landcover); ssf109 <- rename(ssf109, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf109 <- ssf109 %>% extract_covariates(treecover); ssf109 <- rename(ssf109, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf109 <- ssf109 %>% extract_covariates(topo); ssf109 <- rename(ssf109, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf109 <- ssf109 %>% extract_covariates(water); ssf109 <- rename(ssf109, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf109, width=Inf)
```


Jaguar 110
Let's illustrate track regularization with ID = 110. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp110<-trk %>% filter(id=="110") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp110 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp110)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf110 <- temp110 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf110 <- ssf110 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf110 <- ssf110 %>% extract_covariates(dist2drainage); ssf110 <- rename(ssf110, dist2drainage = UTM_b115_dist2drainage_exp)
ssf110 <- ssf110 %>% extract_covariates(dist2water); ssf110 <- rename(ssf110, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf110 <- ssf110 %>% extract_covariates(dist2forest);ssf110 <- rename(ssf110, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf110 <- ssf110 %>% extract_covariates(dist2roads); ssf110 <- rename(ssf110, dist2roads = UTM_b115_dist2roads_exp) 
ssf110 <- ssf110 %>% extract_covariates(human_footprint); ssf110 <- rename(ssf110, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf110 <- ssf110 %>% extract_covariates(Livestock); ssf110 <- rename(ssf110, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf110 <- ssf110 %>% extract_covariates(popdensity); ssf110 <- rename(ssf110, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf110 <- ssf110 %>% extract_covariates(Landcover); ssf110 <- rename(ssf110, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf110 <- ssf110 %>% extract_covariates(treecover); ssf110 <- rename(ssf110, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf110 <- ssf110 %>% extract_covariates(topo); ssf110 <- rename(ssf110, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf110 <- ssf110 %>% extract_covariates(water); ssf110 <- rename(ssf110, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf110, width=Inf)
```


Jaguar 111
Let's illustrate track regularization with ID = 111. Let's
keeping it as 3 hours with tolerance of 1 hours

```{r }
temp111<-trk %>% filter(id=="111") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp111 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp111)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf111 <- temp111 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf111 <- ssf111 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf111 <- ssf111 %>% extract_covariates(dist2drainage); ssf111 <- rename(ssf111, dist2drainage = UTM_b115_dist2drainage_exp)
ssf111 <- ssf111 %>% extract_covariates(dist2water); ssf111 <- rename(ssf111, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf111 <- ssf111 %>% extract_covariates(dist2forest);ssf111 <- rename(ssf111, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf111 <- ssf111 %>% extract_covariates(dist2roads); ssf111 <- rename(ssf111, dist2roads = UTM_b115_dist2roads_exp) 
ssf111 <- ssf111 %>% extract_covariates(human_footprint); ssf111 <- rename(ssf111, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf111 <- ssf111 %>% extract_covariates(Livestock); ssf111 <- rename(ssf111, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf111 <- ssf111 %>% extract_covariates(popdensity); ssf111 <- rename(ssf111, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf111 <- ssf111 %>% extract_covariates(Landcover); ssf111 <- rename(ssf111, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf111 <- ssf111 %>% extract_covariates(treecover); ssf111 <- rename(ssf111, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf111 <- ssf111 %>% extract_covariates(topo); ssf111 <- rename(ssf111, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf111 <- ssf111 %>% extract_covariates(water); ssf111 <- rename(ssf111, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf111, width=Inf)
```


Jaguar 112
Let's illustrate track regularization with ID = 112. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp112<-trk %>% filter(id=="112") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp112 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp112)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf112 <- temp112 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf112 <- ssf112 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf112 <- ssf112 %>% extract_covariates(dist2drainage); ssf112 <- rename(ssf112, dist2drainage = UTM_b115_dist2drainage_exp)
ssf112 <- ssf112 %>% extract_covariates(dist2water); ssf112 <- rename(ssf112, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf112 <- ssf112 %>% extract_covariates(dist2forest);ssf112 <- rename(ssf112, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf112 <- ssf112 %>% extract_covariates(dist2roads); ssf112 <- rename(ssf112, dist2roads = UTM_b115_dist2roads_exp) 
ssf112 <- ssf112 %>% extract_covariates(human_footprint); ssf112 <- rename(ssf112, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf112 <- ssf112 %>% extract_covariates(Livestock); ssf112 <- rename(ssf112, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf112 <- ssf112 %>% extract_covariates(popdensity); ssf112 <- rename(ssf112, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf112 <- ssf112 %>% extract_covariates(Landcover); ssf112 <- rename(ssf112, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf112 <- ssf112 %>% extract_covariates(treecover); ssf112 <- rename(ssf112, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf112 <- ssf112 %>% extract_covariates(topo); ssf112 <- rename(ssf112, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf112 <- ssf112 %>% extract_covariates(water); ssf112 <- rename(ssf112, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf112, width=Inf)
```


Jaguar 113
Let's illustrate track regularization with ID = 113. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp113<-trk %>% filter(id=="113") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp113 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp113)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf113 <- temp113 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf113 <- ssf113 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf113 <- ssf113 %>% extract_covariates(dist2drainage); ssf113 <- rename(ssf113, dist2drainage = UTM_b115_dist2drainage_exp)
ssf113 <- ssf113 %>% extract_covariates(dist2water); ssf113 <- rename(ssf113, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf113 <- ssf113 %>% extract_covariates(dist2forest);ssf113 <- rename(ssf113, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf113 <- ssf113 %>% extract_covariates(dist2roads); ssf113 <- rename(ssf113, dist2roads = UTM_b115_dist2roads_exp) 
ssf113 <- ssf113 %>% extract_covariates(human_footprint); ssf113 <- rename(ssf113, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf113 <- ssf113 %>% extract_covariates(Livestock); ssf113 <- rename(ssf113, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf113 <- ssf113 %>% extract_covariates(popdensity); ssf113 <- rename(ssf113, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf113 <- ssf113 %>% extract_covariates(Landcover); ssf113 <- rename(ssf113, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf113 <- ssf113 %>% extract_covariates(treecover); ssf113 <- rename(ssf113, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf113 <- ssf113 %>% extract_covariates(topo); ssf113 <- rename(ssf113, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf113 <- ssf113 %>% extract_covariates(water); ssf113 <- rename(ssf113, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf113, width=Inf)
```


Jaguar 114
Let's illustrate track regularization with ID = 114. Let's
keeping is as 3 hours with tolerance of 1 hours

```{r }
temp114<-trk %>% filter(id=="114") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp114 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp114)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf114 <- temp114 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf114 <- ssf114 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf114 <- ssf114 %>% extract_covariates(dist2drainage); ssf114 <- rename(ssf114, dist2drainage = UTM_b115_dist2drainage_exp)
ssf114 <- ssf114 %>% extract_covariates(dist2water); ssf114 <- rename(ssf114, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf114 <- ssf114 %>% extract_covariates(dist2forest);ssf114 <- rename(ssf114, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf114 <- ssf114 %>% extract_covariates(dist2roads); ssf114 <- rename(ssf114, dist2roads = UTM_b115_dist2roads_exp) 
ssf114 <- ssf114 %>% extract_covariates(human_footprint); ssf114 <- rename(ssf114, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf114 <- ssf114 %>% extract_covariates(Livestock); ssf114 <- rename(ssf114, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf114 <- ssf114 %>% extract_covariates(popdensity); ssf114 <- rename(ssf114, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf114 <- ssf114 %>% extract_covariates(Landcover); ssf114 <- rename(ssf114, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf114 <- ssf114 %>% extract_covariates(treecover); ssf114 <- rename(ssf114, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf114 <- ssf114 %>% extract_covariates(topo); ssf114 <- rename(ssf114, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf114 <- ssf114 %>% extract_covariates(water); ssf114 <- rename(ssf114, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf114, width=Inf)
```


Jaguar 115
Let's illustrate track regularization with ID = 115. Let's
keeping it as 3 hours with tolerance of 1 hours

```{r }
temp115<-trk %>% filter(id=="115") %>% track_resample(rate=hours(3), tolerance=hours(1))
temp115 %>% select(id, x_, y_, t_, burst_)

#Just checking it
summarize_sampling_rate(temp115)
```

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function steps_by_burst, which in contrast to the steps function accounts for bursts.

```{r }
ssf115 <- temp115 %>% steps_by_burst()
```

Next, we generate random steps with the function random_steps. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with n random steps.

```{r }
ssf115 <- ssf115 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with extract_covariates.

```{r }
ssf115 <- ssf115 %>% extract_covariates(dist2drainage); ssf115 <- rename(ssf115, dist2drainage = UTM_b115_dist2drainage_exp)
ssf115 <- ssf115 %>% extract_covariates(dist2water); ssf115 <- rename(ssf115, dist2water = UTM_b115_dist2waterbodies_exp) 
ssf115 <- ssf115 %>% extract_covariates(dist2forest);ssf115 <- rename(ssf115, dist2forest =UTM_b115_dist2forestedges_Neotropic_Hansen_forest1_0_95percenttreecover_2000_30_tif_exp_edge_dist)  
ssf115 <- ssf115 %>% extract_covariates(dist2roads); ssf115 <- rename(ssf115, dist2roads = UTM_b115_dist2roads_exp) 
ssf115 <- ssf115 %>% extract_covariates(human_footprint); ssf115 <- rename(ssf115, human_footprint = UTM_b115_human_footprint_2009_1km_tif_exp) 
ssf115 <- ssf115 %>% extract_covariates(Livestock); ssf115 <- rename(ssf115, Livestock = UTM_b115_Livestock_Cattle_CC2006_AD_1km_neotropic_albers_tif_exp) 
ssf115 <- ssf115 %>% extract_covariates(popdensity); ssf115 <- rename(ssf115, popdensity =UTM_b115_Population_density_gpw_v4_rev10_2015_1km_neotropic_albers_tif_exp) 
ssf115 <- ssf115 %>% extract_covariates(Landcover); ssf115 <- rename(ssf115, Landcover = UTM_b115_Landcover_ESACCI_2015_300m_neotropic_albers_tif_exp) 
ssf115 <- ssf115 %>% extract_covariates(treecover); ssf115 <- rename(ssf115, treecover = UTM_b115_Neotropic_Hansen_percenttreecover_2000_30m_tif_exp) 
ssf115 <- ssf115 %>% extract_covariates(topo); ssf115 <- rename(ssf115, topo = UTM_b115_Neotropic_Earthenv_dem90m_tif_exp)  
ssf115 <- ssf115 %>% extract_covariates(water); ssf115 <- rename(ssf115, water = UTM_b115_water_frequency_2010_30m_tif_exp)  

print(ssf115, width=Inf)

# Organize variables
#numeric
ssf105$case_<-as.numeric(ssf105$case_)
```

landcover classes 

```{r }
ssf105$landcover <-as.factor(ssf105$Landcover)
ssf105<-ssf105 %>% mutate(landcover = fct_collapse(landcover,
                                                crop = c("10", "20"),
                                                mosaic_crop = c("30"),
                                                forest =c("60","70","80"),
                                                mosaic_treeshrub =c("100"),
                                                mosaic_herb =c("110"),
                                                sparse_tsh =c("150"),
                                                flood_tree =c("160","170"),
                                                flood_shrubherb = c("180"),
                                                shrub= c("120"),
                                                grass = c("130"),
                                                water= c("210"),
                                                urban = c("190"),
                                                bare = c("200")))
print(ssf105, width=Inf)
```

Center and scale variables

```{r }
ssf105<-ssf105 %>% mutate(tcoverSC=as.numeric(scale(treecover)),livestockSC=as.numeric(scale(Livestock)),
                          topoSC=as.numeric(scale(topo)),
                          humanfootSC=as.numeric(scale(human_footprint)), waterSC=as.numeric(scale(water)), 
                          dist2forestSC=as.numeric(scale(dist2forest)),dist2roadSC=as.numeric(scale(dist2roads)),
                          dist2waterSC=as.numeric(scale(dist2water)),distdrainSC=as.numeric(scale(dist2drainage)),
                          popSC=as.numeric(scale(popdensity))) 
```

## Explore the data:
Look Distribution of variables for used and available

```{r fig.width=8, fig.height=8}
binomial_smooth <- function(...) {
  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
}
ggplot(ssf105,aes(x=elevat, y=case_))+
stat_smooth(method="glm", method.args = list(family = "binomial"))+
  binomial_smooth(formula = y ~ splines::ns(x, 5), colour="red")+
  facet_wrap(~id, scales="free")
```

ALTERNATIVELY
Now loop over individuals and do the following:

- Regularize trajectories using an appropriate time window (see e.g., below) 
- calculate new dt values
- Create bursts using individual-specific time intervals
- Generate random steps within each burst

The random steps are generated using the following approach:

1. Fit a gamma distribution to step lenghts
2. Fit a von mises distribution to turn angles
3. Use these distribution to draw new turns and step lengths, form new simulated steps
and generate random x,y values.


```{r }
ssfdat<-NULL
temptrk<-with(trk, track(x=x_, y=y_, t=t_, id=id)) ###,Event_ID=Event_ID,project_region=project_region, sex=sex, age=age, weight=weight,status=status, period=period))
uid<-unique(trk$id) # individual identifiers
luid<-length(uid) # number of unique individuals
for(i in 1:luid){
  # Subset individuals & regularize track 
  temp<-temptrk%>% filter(id==uid[i]) %>% 
  track_resample(rate=hours(round(timestats$median[i])), 
                   tolerance=hours(max(3,round(timestats$median[i]/6))))
				   
  #temp<-temptrk%>% filter(id==uid[i]) %>% 
    #track_resample(rate=hours(8), tolerance=hours(4)) 
### This type of approach would discard many points and increase the median time differences
	
  # Get rid of any bursts without at least 3 points
  temp<-filter_min_n_burst(temp, min_n =3) 
  
  # burst steps
  stepstemp<-steps_by_burst(temp)
  
  # create random steps using fitted gamma and von mises distributions and append
  rnd_stps <- stepstemp %>% random_steps(n = 15)
  
###Here we may include the adjust_param functions from amt package
###(adjust_shape, adjust_scale, adjust_scale, adjust_kappa) to create a iSSF
  
  # append id
  rnd_stps<-rnd_stps%>%mutate(id=uid[i])
  
  # append new data to data from other individuals
  ssfdat<-rbind(rnd_stps, ssfdat)
}
ssfdat<-as_tibble(ssfdat)
ssfdat

###Just to check the medians of what we did in comparison with the totals
ssfdatemp<-ssfdat %>% filter(case_==TRUE) 
ssfdatdf <- data.frame(ssfdatemp)
# A summary applied to ungrouped tbl returns a single row group by id,burst_
bursts <-ssfdatdf %>%
group_by(id,burst_) %>%
summarise(n_distinct = n_distinct(burst_), n = n())%>% data.frame
head(bursts)  

  	
#SSFs sampled data (x1, y1, t1)
#ssfdattrk <- mk_track(ssfdatemp, .x=x1_, .y=y1_, .t=t1_, id=id, crs = CRS("+proj=utm +zone=21K +south +datum=WGS84 +units=m +no_defs")) 
##Summary of totals from the original trk(as we did above just to compare)
#(timestatsssf<-ssfdattrk %>% nest(-id) %>% mutate(sr = map(data, summarize_sampling_rate, time_unit = "hour")) %>% select(id, sr) %>% unnest)
### SSFs sampled data (x2, y2, t2)
#ssfdattrk <- mk_track(ssfdatemp, .x=x2_, .y=y2_, .t=t2_, id=id, crs = CRS("+proj=utm +zone=21K +south +datum=WGS84 +units=m +no_defs")) 
##Summary of totals from the original trk(as we did above just to compare)
#(timestatsssf<-ssfdattrk %>% nest(-id) %>% mutate(sr = map(data, summarize_sampling_rate, time_unit = "hour")) %>% select(id, sr) %>% unnest)
## And comparing that to the original file
#SaoBentotrk -> trk	
#(timestats<-trk %>% nest(-id,-sex,-age,-weight,-status) %>% mutate(sr = map(data, summarize_sampling_rate, time_unit = "hour")) %>%select(id,sex,age,weight,status,sr) %>% unnest)
```

Now, lets plot the data for random and matched points


```{r fig.height=12, fig.width=12, warning=FALSE}
ggplot(ssfdat, aes(x2_, y2_, color=case_))+geom_point()+facet_wrap(~id,scales="free")
```

## Write out data for further annotating
 With the SSFs, we have the extra complication of
having a time and location at both the start and end of the step.  

For the time being, we will assume we want to annotate variables at the end of the step
but use the starting point of the step as the timestamp.

```{r }
ssf.df <- data.frame(ssfdat)
head(ssf.df)
ssf.df$timestamp <-ssf.df$t1_
ssf.df$x <-ssf.df$x2_
ssf.df$y <-ssf.df$y2_
head(ssf.df)
ssfALL<-ssf.df
head(ssfALL)
#ssfALL$sl_ <- NULL; ssfALL$ta_<- NULL
ssfALL<-ssfALL%>% mutate(obs = 1:n()) %>% select(obs, everything())
head(ssfALL)
#stepSB<-ssfALL %>% select("obs","id","case_", "x", "y","timestamp","step_id_")
#head(stepSB)
#AllSteps <- ssf.df  ### include name of the area/ project
```

These points then need to be annotated prior to fitting ssfs. 2 files All and to annotate

```{r }
#write.csv(ssfALL, file="C:/RWorkDir/jaguardatapaper/ssfALL.csv", row.names = FALSE) ###annotated_projectname
#write.csv(stepSB, file="C:/RWorkDir/jaguardatapaper/stepSB.csv", row.names = FALSE) ###annotated_projectname
```



```{r }
#" Simplify some variable names and make case a numeric variable
ssfdat$case_<-as.numeric(ssfdat$case)
```

landcover classes 

```{r }
ssfdat$land_cover <-as.factor(ssfdat$land_cover)


ssfdat<-ssfdat %>% mutate(land_C = fct_collapse(land_use,
                                                crop = c("10", "20"),
                                                mosaic_crop = c("30"),
                                                forest =c("60","70","80"),
                                                mosaic_treeshrub =c("100"),
                                                mosaic_herb =c("110"),
                                                sparse_tsh =c("150"),
                                                flood_tree =c("160","170"),
                                                flood_shrubherb = c("180"),
                                                shrub= c("120"),
                                                grass = c("130"),
                                                water= c("210"),
                                                urban = c("190"),
                                                bare = c("200")))
```

